---
title: "eda"
author: "Sophia Gan"
date: "2024-04-08"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data <- read.csv("train_data.csv")
```

```{r}
library(corrplot)
library(caret)
library(ggplot2)
library(glmnet)
library(Matrix)
library(ggfortify)
```

```{r}
summary(data)
str(data)
head(data)
```

## except for AspectRation, ShapeFactor3, Compactness, others might need normalization
```{r}
for (variable in names(data)) {
  plot <- ggplot(data, aes_string(x = variable)) +
          geom_density(fill = "blue", alpha = 0.7) +
          ggtitle(paste("Density of", variable)) +
          xlab(variable) +
          ylab("Density")
  
  print(plot)
}
```

## correlation on the original dataset
```{r}
correlation_matrix <- cor(data)
corrplot(correlation_matrix, method = "circle")
```
## reducing the data by removing the highly correlated ones, discuss on the cutoff
## not sure which data should be standardized or normalized, initial guess is 
```{r}
highCor <- findCorrelation(correlation_matrix, cutoff = 0.75)
data_reduced <- data[,-highCor]
```

## standardize before lasso/ridge and pca
```{r}
numeric_columns <- sapply(data, is.numeric)
numeric_columns["Y"] <- FALSE
numeric_columns
data[numeric_columns] <- scale(data[numeric_columns])
```

## lasso on standardized dataset
```{r}
library(glmnet)
x <- as.matrix(data[, -which(names(data) == "Y")])
y <- data$Y
cv_fit <- cv.glmnet(x, y, alpha = 1) #alpha=1 for lasso
best_lambda <- cv_fit$lambda.min
best_lambda
modeling <- glmnet(x, y, family = "gaussian", alpha = 1, lambda = best_lambda)
coef <- coef(modeling)
coef
plot(cv_fit$glmnet.fit, "lambda", label = "TRUE")
coef_table <- as.data.frame.matrix(coef)
coef_table
```

## pca on standardized dataset
```{r}
library("FactoMineR")
pca_result <- prcomp(data, center = TRUE, scale. = TRUE)
summary(pca_result)
autoplot(pca_result, data = data, color = 'Y')  # If 'Y' is the target variable
plot(pca_result)
pca_data <- as.data.frame(pca_result$x)

# normalized <- scale(data)
# data.pca <- princomp(cor(normalized))
# fviz_eig(data.pca, addlabels = TRUE)
```

```{r}
data$Y <- factor(data$Y)
ols.model <- glm(Y ~ ., family = "binomial", data = data)
summary(ols.model)
```

